---
title: "Maternal Education Differences in Vocabulary Development"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: tango
    theme: spacelab
---

```{r knitr, include=FALSE}
library(knitr)
opts_chunk$set(message = FALSE, warning = FALSE)
```

Load required libraries.
```{r packages}
library(wordbankr)
library(langcog)
library(dplyr)
library(ggplot2)
library(directlabels)
theme_set(theme_mikabr())
font <- theme_mikabr()$text$family
mode <- "local"
```

Get administration data and filter to administrations of Words & Sentences that have momed coded.
```{r vocab_admins}
vocab_admins <- get_administration_data(mode = mode) %>%
  select(data_id, language, form, age, mom_ed, production) %>%
  rename(momed = mom_ed) %>%
  filter(form == "WS", !is.na(momed))
```

Get item information to find the number of items on each language's form.
```{r num_words}
num_words <- get_item_data(mode = mode) %>%
  filter(form == "WS", type == "word") %>%
  group_by(language) %>%
  summarise(n = n())
```

Normalize productive vocabulary size as a proportion of items and calculate median vocabulary size for each language, momed level, and age.
```{r vocab_data}
vocab_data <- vocab_admins %>%
  left_join(num_words) %>%
  mutate(production = as.numeric(production) / n,
         momed = `levels<-`(momed, list("None, Primary, Some Secondary" = "None",
                                        "None, Primary, Some Secondary" = "Primary",
                                        "None, Primary, Some Secondary" = "Some Secondary",
                                        "Secondary" = "Secondary",
                                        "Some College" = "Some College",
                                        "College" = "College",
                                        "Some Graduate" = "Some Graduate",
                                        "Graduate" = "Graduate")))

overall_vocab_data <- vocab_data %>%
  group_by(language, momed, age) %>%
  summarise(median = median(production))

cuts <- seq(0, 1, by = 0.25)
percentile_vocab_data <- vocab_data %>%
  group_by(language, momed, age) %>%
  mutate(percentile = rank(production) / length(production),
         quantile = cut(percentile, breaks = cuts))
```

Plot vocabulary size over age by momed.
```{r plot_overall, fig.width=5, fig.height=9}
ggplot(filter(overall_vocab_data, language != "Hebrew"),
       aes(x = age, y = median, colour = momed, label = momed)) +
  facet_grid(language ~ .) +
  geom_jitter(size = 0.7) +
  geom_smooth(se = FALSE, span = 1, method = "loess", size = 1) +
  scale_colour_solarized(name = "") +
  scale_x_continuous(breaks = seq(min(vocab_data$age), max(vocab_data$age), 2),
                     limits = c(min(vocab_data$age), max(vocab_data$age) + 1),
                     name = "\nAge (months)") +
  scale_y_continuous(name = "Median Productive Vocabulary (proportion of total words)\n",
                     limits = c(0,1)) +
  theme(legend.position = "bottom",
        legend.direction = "vertical")
```

Plot vocabulary size over age by momed and percentile.
```{r plot_percentile, fig.width=12, fig.height=10}
ggplot(filter(percentile_vocab_data, language != "Hebrew"),
       aes(x = age, y = production, colour = momed, label = momed)) +
  facet_grid(language ~ quantile) +
  geom_jitter(size = 0.7, alpha = 0.7) +
  geom_smooth(se = FALSE, span = 1, method = "loess", size = 1) +
  scale_colour_solarized(name = "") +
  scale_x_continuous(breaks = seq(min(vocab_data$age), max(vocab_data$age), 2),
                     limits = c(min(vocab_data$age), max(vocab_data$age) + 1),
                     name = "\nAge (months)") +
  scale_y_continuous(name = "Productive Vocabulary (proportion of total words)\n",
                     limits = c(0,1)) +
  theme(legend.position = "bottom")
```
